---
title: "Data Labelling"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data-labeling}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
library(mestrado)
library(dplyr)
```

## Annotations dataset

The function `tidy_annotations()` is just a wrapper around `purrr::map_dfr()` for stack up all the rds files generated by `wavesurfer::annotator_app()`. (see `vignette("04-data-annotation")`)

```{r}
annotations_dir <-  system.file("annotations", package = "mestrado")
annotations_dir

annotations <- tidy_annotations(annotations_dir)
```


```{r}
glimpse(annotations)
annotations %>% head() %>% knitr::kable()
```

## Retriving the labels for the slices

By merging the slices of the wave files with the labels annotated in the previous step, we construct the final dataset with the response/targets ready for modelling. The function `label_slices()` take the directory of slices and the annotations dataset as inputs. The output is a map between slice and label.

```{r}
slices_dir <-  system.file("wav_sample_slices_1000ms", package = "mestrado")

slices_1000ms_labels <- label_slices(
  slices_dir, 
  annotations, 
  pattern = "Glaucidium|Megascops-atricapilla"
)
```

```{r}
glimpse(slices_1000ms_labels)
slices_1000ms_labels %>% head() %>% knitr::kable()
```


