---
title: "Slicing Wave Files"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{data-slicing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The observations of the training set will be slices with a given length of the raw wave files. For instance, if an original wave file has duration of 55 seconds, then the slicing with interval of 1 second and no overlap will result in 55 disjoint 1 second long slices.

```{r}
library(mestrado)

wav_dir <- system.file("wav_sample", package = "mestrado")
temp_dir <- tempdir()

slices_path <- slice_wavs(wav_dir, temp_dir)
slices_path

slices <- list.files(slices_path)
slices[4:7]
```

The resulting file names was designed to make it "parser friendly". It goes well with `tidyr::separate(sep = "@")`. This data wis useful when matching with the annotations of the presense/absensce of a bird song or any type of event of interest.

```{r, message = FALSE, warning=FALSE}
library(tidyverse)
slices_metadata <- tibble(
  file_name = slices
) %>%
  tidyr::separate(file_name, c("species", "start", "end"), sep = "@")

slices_metadata %>% head()
```

