--- 
title: "Mestrado"
author: "Athos Petri Damiani"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::pdf_book
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: athospd/mestrado
description: "Documentação do Mestrado"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(stats)
library(readr)
library(purrr)
library(slickR)
library(dplyr)
library(ggplot2)
library(forcats)
```


# Projeto

Projeto de pesquisa *"Detectores acústicos automáticos de espécies de aves noturnas em seu bioma natural"* do Departamento de Engenharia Mecânica da Escola Politécnica (Laboratório de Acústica e Meio Ambiente).

**Orientador:** Prof. Dr. Linilson Padovese

**Co-orientador:** Dr. Paulo Hubert

## Escopo


## Cronograma


## Notas

> "There is the need for shared datasets with annotations of a wide variety of calls for a large number of species if methods that are suitable for conservation work are to be developed."
> --- Automated birdsong recognition in complex acoustic environments: a review
